On selection of a coupling method to cope with change of priors in one-vs-one classification

### 2025-08-14 PatRec.tex
Correct hyphenation of Wu-Lin-Weng

\documentclass[twoside,11pt]{article}

@@

\usepackage{amsmath, amsthm, amssymb}
\usepackage{jmlr2e}

@@

\usepackage{lastpage}
\jmlrheading{23}{2025}{1-\pageref{LastPage}}{1/21; Revised 5/22}{9/22}{21-0000}{Author One and Author Two}

% Short headings should be running head and authors last names

\ShortHeadings{Probabilistic OVO}{Probabilistic OVO}
\firstpageno{1}

@@

\title{New Bayes covariant coupling methods to cope with change of priors in one-vs-one classification }

\author{\name Ondrej Šuch \email ondrejs@savbb.sk \\
       \addr Matematický ústav SAV\\
       Ďumbierska 1\\
       Banská Bystrica, 974 01, Slovakia
       \AND
       \name Peter Novotný \email peter.novotny@fri.uniza.sk \\
       \addr Fakulta informatiky a riadenia\\
       Žilinská Univerzita v Žiline\\
       Žilina, 010 26, Slovakia
       \AND
       \name Ali Haidar \email haidar@savbb.sk \\
       \begin{minipage}[t]{0.45\textwidth}
       \addr Matematický ústav SAV\\
       Ďumbierska 1\\
       Banská Bystrica, 974 01, Slovakia
       \end{minipage}\hfill
       \begin{minipage}[t]{0.45\textwidth}
       \addr Fakulta matematiky, fyziky a informatiky\\
       Univerzity Komenského \\
       Mlynská dolina F1 \\
       Bratislava, 842 48, Slovakia
       \end{minipage}
       }

\editor{My editor}

@@

\begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
%\blindtext
Label shift is one kind of distribution shift affecting classification models. Generally, one uses transformation inspired by Bayes classifiers to 
adjust predictions of probabilistic classification models. However, there is an ambiguity how to apply this transformation in the case of  probabilistic models built in one-vs-one (also called pairwise) fashion. One-vs-one classification relies on the choice of a coupling method. Some coupling methods -- called Bayes covariant --  are not affected by the ambiguity. Adopting a Bayes covariant method would be the ideal solution to the ambiguity. However, the most commonly used coupling  method of Wu-Lin-Weng is not Bayes covariant. In our study we present elements of theory of coupling methods and exhibit new Bayes covariant coupling methods. One class is formed by arboreal methods. Still more methods can be formed by the way of ensembling. We compare their performance with that of Wu-Lin-Weng's coupling method. In majority of cases, it is possible to find a Bayes covariant method matching performance of Wu-Lin-Weng's method.

\end{abstract}

\begin{keywords}
  distribution shift, label shift, pairwise coupling, one-vs-one classification, Bayes classifier
\end{keywords}

>>>>

%\documentclass[twoside,11pt]{article}
\documentclass[a4paper,fleqn]{cas-sc}

\usepackage[authoryear,longnamesfirst]{natbib}

@@

\usepackage{amsmath, amsthm, amssymb}
%\usepackage{jmlr2e}

@@

\usepackage{lastpage}
%\jmlrheading{23}{2025}{1-\pageref{LastPage}}{1/21; Revised 5/22}{9/22}{21-0000}{Author One and Author Two}
%
%% Short headings should be running head and authors last names
%
%\ShortHeadings{Probabilistic OVO}{Probabilistic OVO}
%\firstpageno{1}

%%%Author macros
\def\tsc#1{\csdef{#1}{\textsc{\lowercase{#1}}\xspace}}
\tsc{WGM}
\tsc{QE}
\tsc{EP}
\tsc{PMS}
\tsc{BEC}
\tsc{DE}
%%%

@@

\let\WriteBookmarks\relax
\def\floatpagepagefraction{1}
\def\textpagefraction{.001}
\shorttitle{Leveraging social media news}
\shortauthors{J.K. Krishnan et~al.}
%\begin{frontmatter}

%	\title [mode = title]{This is a specimen $a_b$ title}                      
\title{New Bayes covariant coupling methods to cope with change of priors in one-vs-one classification }
\tnotemark[1,2]

\tnotetext[1]{This document is the results of the research
		project funded by the National Science Foundation.}

\tnotetext[2]{The second title footnote which is a longer text matter
		to fill through the whole text width and overflow into
		another line in the footnotes area of the first page.}

\author[1]{O. Šuch}
\cormark[1]
\ead{ondrejs@savbb.sk}

\affiliation[1]{organization={Matematický ústav Slovenskej Akadémie Vied},
		addressline={Ďumbierska 1},
		city = {Banská Bystrica},
		postcode = {97411},
		country = {Slovakia}}

\author[2]{P. Novotný}
\cormark[2]
\ead{peter.novotny@fri.uniza.sk}

\affiliation[2]{organization = {Žilinská Univerzita v Žiline},
addressline={Univerzitná 8215/1},
city={Žilina},
postcode = {01026},
country = {Slovakia}}

%\author{\name Ondrej Šuch \email ondrejs@savbb.sk \\
%       \addr Matematický ústav SAV\\
%       Ďumbierska 1\\
%       Banská Bystrica, 974 01, Slovakia
%       \AND
%       \name Peter Novotný \email peter.novotny@fri.uniza.sk \\
%       \addr Fakulta informatiky a riadenia\\
%       Žilinská Univerzita v Žiline\\
%       Žilina, 010 26, Slovakia
%       \AND
%       \name Ali Haidar \email haidar@savbb.sk \\
%       \begin{minipage}[t]{0.45\textwidth}
%       \addr Matematický ústav SAV\\
%       Ďumbierska 1\\
%       Banská Bystrica, 974 01, Slovakia
%       \end{minipage}\hfill
%       \begin{minipage}[t]{0.45\textwidth}
%       \addr Fakulta matematiky, fyziky a informatiky\\
%       Univerzity Komenského \\
%       Mlynská dolina F1 \\
%       Bratislava, 842 48, Slovakia
%       \end{minipage}
%       }

%\editor{My editor}

\begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
%\blindtext
Label shift is one kind of distribution shift affecting classification models. Generally, one uses transformation inspired by Bayes classifiers to 
adjust predictions of probabilistic classification models. However, there is an ambiguity how to apply this transformation in the case of  probabilistic models built in one-vs-one (also called pairwise) fashion. One-vs-one classification relies on the choice of a coupling method. Some coupling methods -- called Bayes covariant --  are not affected by the ambiguity. Adopting a Bayes covariant method would be the ideal solution to the ambiguity. However, the most commonly used coupling  method of Wu-Lin-Weng is not Bayes covariant. In our study we present elements of theory of coupling methods and exhibit new Bayes covariant coupling methods. One class is formed by arboreal methods. Still more methods can be formed by the way of ensembling. We compare their performance with that of Wu-Lin-Weng's coupling method. In majority of cases, it is possible to find a Bayes covariant method matching performance of Wu-Lin-Weng's method. 
\end{abstract}

\begin{highlights}
\item New Bayes covariant coupling methods
\item New way to create ensembles of multiple Bayes covariant methods
\item Unification of Hastie-Tibshirani and Wu-Lin-Weng method
\item Rigorous evaluation of coupling methods
\end{highlights}

\begin{keywords}
  distribution shift \sep label shift \sep pairwise coupling \sep one-vs-one classification \sep Bayes classifier
\end{keywords}

@@

### 2025-08-14 coupling.tex
Correct hyphenation of Wu-Lin-Weng

The quadratic score is the basis of the popular Wu-Lin-Weng's method \cite{wu2004probability}. 
Their method uses an additional trick, which we term \emph{self-referentiality}. In the method of Hastie and Tibshirani, $w_{ij}$ are constants derived from observed class frequencies. However, we could use weights $w_{ij}$ that increase when the predicted probability of either class $i$ or $j$ is high. A reasonable choice would be for instance $w_{ij}= p_i + p_j$. In the case of quadratic score, it is more convenient to use weights $w_{ij} = (p_i + p_j)^2$ because then the resulting optimization problem is just minimization of a quadratic form
\begin{align*}
\hat{\boldsymbol{p}} \stackrel{def}{=} \argmin_{\boldsymbol{p}} \sum_{i,j} (r_{ij}p_j - r_{ji}p_i)^2.
\end{align*}

We note that the method is symmetric. The choice of quadratic scoring rule can also be viewed as canonical in view of axiomatic characterization given in  \cite{selten1998axiomatic}.  However the choice of weights $w_{ij}$ to make the method computationally amenable is somewhat arbitrary. Thus we do not view Wu-Lin-Weng's method as canonical.

>>>>

The quadratic score is the basis of the popular Wu--Lin--Weng's method \cite{wu2004probability}. 
Their method uses an additional trick, which we term \emph{self-referentiality}. In the method of Hastie and Tibshirani, $w_{ij}$ are constants derived from observed class frequencies. However, we could use weights $w_{ij}$ that increase when the predicted probability of either class $i$ or $j$ is high. A reasonable choice would be for instance $w_{ij}= p_i + p_j$. In the case of quadratic score, it is more convenient to use weights $w_{ij} = (p_i + p_j)^2$ because then the resulting optimization problem is just minimization of a quadratic form
\begin{align*}
\hat{\boldsymbol{p}} \stackrel{def}{=} \argmin_{\boldsymbol{p}} \sum_{i,j} (r_{ij}p_j - r_{ji}p_i)^2.
\end{align*}

We note that the method is symmetric. The choice of quadratic scoring rule can also be viewed as canonical in view of axiomatic characterization given in  \cite{selten1998axiomatic}.  However the choice of weights $w_{ij}$ to make the method computationally amenable is somewhat arbitrary. Thus we do not view Wu--Lin--Weng's method as canonical.

### 2025-08-14 discussion.tex
Correct hyphenation of Wu-Lin-Weng

\begin{itemize}
	\item The orthodox method of Wu--Lin--Weng dominates others across diverse classification tasks.
	\item The only method close to Wu--Lin--Weng is the radial coupling method.
	\item The concern of Hinton about one-vs-one classification is unwarranted. The method that ignores binary classifiers not trained on a given sample (we call it Hinton's oracle) underperforms the method of Wu--Lin--Weng.
	\item In more than half of classification problems we were able to find a Bayes covariant coupling method which matches performance of the Wu-Lin-Weng method. 
\end{itemize}

@@

\begin{table}[!ht]
	\begin{tabular}{cm{2.5cm}|m{1.5cm}ccm{1.5cm}|m{1.5cm}}
	%	&& less competitive & \multicolumn{4}{c|}{competitive method} & oracle \\
		\hline 
		&methods % & Hastie-Tibshirani 
			& Wu--Lin--Weng & radial & normal & arboreal & Hinton's oracle \\
		\hline 
		&source % & \cite{hastie1998classification} 
			& \cite{wu2004probability} & \cite{vsuch2015new} & \cite{vsuch2016bayes} & this work & this work \\
		\multirow{5}{*}{\begin{turn}{90}\makecell{exact}\end{turn}}
		&uniqueness % &  yes 
			& yes & yes & yes & yes & yes \\
		&Bradley--Terry consistency %& yes 
			& yes & yes & yes & yes & yes \\
		&canonicity %& yes 
			& no & yes & yes & no & yes \\
		&symmetry % & yes 
			& yes & yes & yes & no & yes \\
		& Bayes covariant %& no 
			& no & no & yes & yes & yes \\
		\hline
		\multirow{5}{*}{\begin{turn}{90}\makecell{non-exact}\end{turn}}
		&Hinton's minimality %& N/A 
			& \multicolumn{5}{c}{see Section \ref{sec:exp1}} \\
		&robustness %& N/A 
			& \multicolumn{5}{c}{see Section \ref{sec:exp1}} \\
		&lack of parameters %& yes 
			& yes & yes & yes & yes & yes \\
		& amounts to a linear system % & no 
			& yes & yes & yes & yes & yes\\
		& linear system of special form % & N/A 
			& no & yes & yes & yes & yes \\
		& iterative method %& yes 
			& yes & yes & no & not needed & not needed\\
		\hline
	\end{tabular}
	\caption{Summary of properties of coupling methods described in Section \ref{sec:coupling}. The method of Hastie-Tibshirani is omitted from evaluation in Section \ref{sec:exp1}, because earlier evaluation showed it lacks robustness compared to the method of Wu-Lin-Weng \cite{wu2004probability}.}
	\label{tab:summaryCoupling}
\end{table}

@@

Based on our findings we see two immediate pathways for future research. First, one can study arboreal methods in more detail, since these are Bayes covariant and as shown in Section \ref{sec:exp2} they often provide performance matching that of Wu-Lin-Weng's method. Secondly, there is a continuum of Bayes covariant methods which we barely sketched. Their properties, parameter selection,  and applicability to practical tasks need to investigated in greater detail.

>>>>

\begin{itemize}
	\item The orthodox method of Wu--Lin--Weng dominates others across diverse classification tasks.
	\item The only method close to Wu--Lin--Weng is the radial coupling method.
	\item The concern of Hinton about one-vs-one classification is unwarranted. The method that ignores binary classifiers not trained on a given sample (we call it Hinton's oracle) underperforms the method of Wu--Lin--Weng.
	\item In more than half of classification problems we were able to find a Bayes covariant coupling method which matches performance of the Wu--Lin--Weng method. 
\end{itemize}

@@

\begin{table}[!ht]
	\begin{tabular}{cm{2.5cm}|m{1.5cm}ccm{1.5cm}|m{1.5cm}}
	%	&& less competitive & \multicolumn{4}{c|}{competitive method} & oracle \\
		\hline 
		&methods % & Hastie-Tibshirani 
			& Wu--Lin--Weng & radial & normal & arboreal & Hinton's oracle \\
		\hline 
		&source % & \cite{hastie1998classification} 
			& \cite{wu2004probability} & \cite{vsuch2015new} & \cite{vsuch2016bayes} & this work & this work \\
		\multirow{5}{*}{\begin{turn}{90}\makecell{exact}\end{turn}}
		&uniqueness % &  yes 
			& yes & yes & yes & yes & yes \\
		&Bradley--Terry consistency %& yes 
			& yes & yes & yes & yes & yes \\
		&canonicity %& yes 
			& no & yes & yes & no & yes \\
		&symmetry % & yes 
			& yes & yes & yes & no & yes \\
		& Bayes covariant %& no 
			& no & no & yes & yes & yes \\
		\hline
		\multirow{5}{*}{\begin{turn}{90}\makecell{non-exact}\end{turn}}
		&Hinton's minimality %& N/A 
			& \multicolumn{5}{c}{see Section \ref{sec:exp1}} \\
		&robustness %& N/A 
			& \multicolumn{5}{c}{see Section \ref{sec:exp1}} \\
		&lack of parameters %& yes 
			& yes & yes & yes & yes & yes \\
		& amounts to a linear system % & no 
			& yes & yes & yes & yes & yes\\
		& linear system of special form % & N/A 
			& no & yes & yes & yes & yes \\
		& iterative method %& yes 
			& yes & yes & no & not needed & not needed\\
		\hline
	\end{tabular}
	\caption{Summary of properties of coupling methods described in Section \ref{sec:coupling}. The method of Hastie-Tibshirani is omitted from evaluation in Section \ref{sec:exp1}, because earlier evaluation showed it lacks robustness compared to the method of Wu--Lin--Weng \cite{wu2004probability}.}
	\label{tab:summaryCoupling}
\end{table}

@@

Based on our findings we see two immediate pathways for future research. First, one can study arboreal methods in more detail, since these are Bayes covariant and as shown in Section \ref{sec:exp2} they often provide performance matching that of Wu--Lin--Weng's method. Secondly, there is a continuum of Bayes covariant methods which we barely sketched. Their properties, parameter selection,  and applicability to practical tasks need to investigated in greater detail.

### 2025-08-14 experiment.tex
Correct hyphenation of Wu-Lin-Weng

\begin{itemize}
	\item[(a)] Which coupling methods without parameters show good performance across diverse data sets? Is it the orthodox method of Wu-Lin-Weng? An experiment to answer this question is carried out Section \ref{sec:exp1}. 
	\item[(b)] If the answer to a) is a coupling method which is not Bayes covariant, can we find a parametric Bayes covariant method which  matches its performance? An experiment related to this question is carried out in  Sections \ref{sec:exp2} and \ref{sec:exp3}.
	\item[(c)] If the answer to b) is negative, is it better to apply the change of priors before or after coupling? Additionally, how serious is the impact of the lack of Bayes covariance in practice. This question is investigated in Section \ref{sec:exp4}.
\end{itemize}

@@

Small probabilities generally have very little effect on the implementation of Bayes covariant methods which generally operate using log-odds. However small probabilities do affect both the method of Wu-Lin-Weng as well as radial method, because then Assumption \ref{ass:2} fails to hold.

@@

The first experiment compares the performance of the following canonical coupling methods:
\begin{itemize}
\item method of Wu-Lin-Weng, which is widely used in software libraries, and thus can be viewed as an orthodox choice,
\item normal coupling, which is Bayes covariant,
\item radial coupling, which is not Bayes covariant,
\item Hinton's oracle.
\end{itemize}

% \input{tab-multi.tex}
The results are shown in Figure \ref{fig:exp1-plot1}. We can clearly see that Wu-Lin-Weng's method has the most robust performance. On some datasets it trounces normal coupling method by a wide margin. The edge over the radial method is smaller with only mnist and usps datasets showing a significant difference for smaller values of $tol$ hyperparameter.  However in view of Theorem \ref{thm:K3} the Wu-Lin-Weng's method is not Bayes covariant, therefore we will extend the set of Bayes covariant methods beyond the canonical normal coupling.

@@

In this experiment we compare  Bayes covariant coupling methods \emph{without} parameters that could match the performance of Wu-Lin-Weng's method. We investigate  the case of three class datasets. This is a natural starting point, since it is the smallest number of classes where the question is meaningful to ask. All experiments use value of 0.05 for $tol$ hyperparameter based on results of Section \ref{sec:exp1}.  In our comparison we include the normal method, as well as the three arboreal methods $ \boldsymbol{s}_1,~ \boldsymbol{s}_2,~ \boldsymbol{s}_3$ that are not canonical. We consider all possible triples of classes in all datasets. The resulting statistics are shown in Table \ref{tab:step2}.

@@

We can see that in slightly less than half of the cases none of the four parameterless Bayes covariant methods was able to match the performance of Wu-Lin-Weng's method. This leads us to extend evaluation to Bayes covariant methods with parameters.

@@

We start by selecting 1000 three class subsets from all datasets for problems where normal method was outperformed by Wu-Lin-Weng's method. For each three-class classification problem we will try to match the performance of Wu-Lin-Weng method with a Bayes covariant method in family \eqref{eq:family}.

Selecting parameters for methods in family \eqref{eq:family} requires several methodological choices. 
\begin{itemize}
\item To avoid bias in selection of parameters $a_1, a_2, a_3$ we use a subset of testing dataset to find their optimal values. As a result, we will be comparing the accuracy of Wu-Lin-Weng method evaluated with repeated five-fold cross-validation accuracy of a method in family \eqref{eq:family}. We use 20 repetitions.
\item To compare two parameter vectors, we compare their resulting cross-validation accuracy. An alternative approach would be to use  a gradient based optimization for a smooth proxy (e.g. cross-entropy) of empirical accuracy. We opted for the former because the problem is two-dimensional, and we can use grid search instead.
\item Our choice of grid search necessitates a selection of a compact subset of parameter space that will be sampled. We choose to examine points in the simplex defined by  $a_i\geq 0$, for $i=1,2,3$. As Table \ref{tab:step2} shows, this simplex contained in more than half cases a Bayes covariant method matching the performance of Wu-Lin-Weng's method.
\end{itemize}

\begin{figure}[!ht]
\includegraphics{exp2-plot2.pdf}
\caption{Comparison of accuracy for parametric Bayes covariant methods \eqref{eq:family} vs. Wu-Lin-Weng's method}
\label{fig:par-bc}
\end{figure}

The summary statistics are shown in Figure \ref{fig:par-bc}.
We can see that only in minority of cases we found a parametric Bayes covariant coupling method matching the performance of Wu-Lin-Weng's method.

%\subsection{Experiment: }
%
%We determined that among the 1000 three-class problems we have examined, the largest gap between Bayes-covariant parametric methods and Wu-Lin-Weng method occurred for mnist dataset and classes 5,6,9. Figure 
%
%\begin{itemize}
%	\item We searched within a much larger rectangle $D$ containing the simplex $a_i\geq 0$, which in this case is a triangle $T$.
%	\item We used Brier score as a smooth proxy for measuring the quality of multi-class classifier. This score is much smoother than 0-1 score which measures accuracy.
%\end{itemize}
%
%The results are shown in Figure \ref{fig:score}
%
%
%\begin{figure}[!ht]
%	\includegraphics{graphs/exp2-detail.pdf}
%	\caption{Color map indicates the value of Brier score computed for an ensemble of coupling methods on the testing dataset. The vertices triangle indicate the arboreal coupling methods, and the black circle the location of the optimum.}
%	\label{fig:score}
%\end{figure}
%
%We can see that indeed the optimum over the rectangle lies outside of the simplex. Let us compare the resulting accuracy of the optima in $\Delta$ and $D$. The results are shown in Table \ref{tab:step5}.
%
%\input tab-step5.tex

\subsection{Experiment: Effect of change of priors}  \label{sec:exp4}
%\subsection{Wu-Lin-Weng's method}

We have shown that are situations when Wu-Lin-Weng's method outperforms even the best Bayes-covariant parametric methods. In this section we thus ponder the question whether some non-Bayes covariant method would be more resilients to change of priors. We consider two facets to this question.

@@

The results, as expected, detect no change between $\boldsymbol{c}_1, \boldsymbol{c}_2$ for normal coupling method. However, there is no consistent difference in spread between radial and Wu-Lin-Weng's coupling method. Spread for the radial method is lower for 3 datasets (letter, mnist, usps datasets), higher for 3 datasets (dna, satimage, segment) and tied for waveform dataset.

@@

\begin{figure}[!ht]
	\includegraphics{exp4-truth.pdf}
	\caption{Difference in accuracies of $\boldsymbol{c}_2$ and $\boldsymbol{c}_1$ for radial and Wu-Lin-Weng's coupling method.}
	\label{fig:score}
\end{figure}

>>>>

\begin{itemize}
	\item[(a)] Which coupling methods without parameters show good performance across diverse data sets? Is it the orthodox method of Wu--Lin--Weng? An experiment to answer this question is carried out Section \ref{sec:exp1}. 
	\item[(b)] If the answer to a) is a coupling method which is not Bayes covariant, can we find a parametric Bayes covariant method which  matches its performance? An experiment related to this question is carried out in  Sections \ref{sec:exp2} and \ref{sec:exp3}.
	\item[(c)] If the answer to b) is negative, is it better to apply the change of priors before or after coupling? Additionally, how serious is the impact of the lack of Bayes covariance in practice. This question is investigated in Section \ref{sec:exp4}.
\end{itemize}

@@

Small probabilities generally have very little effect on the implementation of Bayes covariant methods which generally operate using log-odds. However small probabilities do affect both the method of Wu--Lin--Weng as well as radial method, because then Assumption \ref{ass:2} fails to hold.

@@

The first experiment compares the performance of the following canonical coupling methods:
\begin{itemize}
\item method of Wu--Lin--Weng, which is widely used in software libraries, and thus can be viewed as an orthodox choice,
\item normal coupling, which is Bayes covariant,
\item radial coupling, which is not Bayes covariant,
\item Hinton's oracle.
\end{itemize}

% \input{tab-multi.tex}
The results are shown in Figure \ref{fig:exp1-plot1}. We can clearly see that Wu--Lin--Weng's method has the most robust performance. On some datasets it trounces normal coupling method by a wide margin. The edge over the radial method is smaller with only mnist and usps datasets showing a significant difference for smaller values of $tol$ hyperparameter.  However in view of Theorem \ref{thm:K3} the Wu--Lin--Weng's method is not Bayes covariant, therefore we will extend the set of Bayes covariant methods beyond the canonical normal coupling.

@@

In this experiment we compare  Bayes covariant coupling methods \emph{without} parameters that could match the performance of Wu--Lin--Weng's method. We investigate  the case of three class datasets. This is a natural starting point, since it is the smallest number of classes where the question is meaningful to ask. All experiments use value of 0.05 for $tol$ hyperparameter based on results of Section \ref{sec:exp1}.  In our comparison we include the normal method, as well as the three arboreal methods $ \boldsymbol{s}_1,~ \boldsymbol{s}_2,~ \boldsymbol{s}_3$ that are not canonical. We consider all possible triples of classes in all datasets. The resulting statistics are shown in Table \ref{tab:step2}.

@@

We can see that in slightly less than half of the cases none of the four parameterless Bayes covariant methods was able to match the performance of Wu--Lin--Weng's method. This leads us to extend evaluation to Bayes covariant methods with parameters.

@@

We start by selecting 1000 three class subsets from all datasets for problems where normal method was outperformed by Wu--Lin--Weng's method. For each three-class classification problem we will try to match the performance of Wu--Lin--Weng method with a Bayes covariant method in family \eqref{eq:family}.

Selecting parameters for methods in family \eqref{eq:family} requires several methodological choices. 
\begin{itemize}
\item To avoid bias in selection of parameters $a_1, a_2, a_3$ we use a subset of testing dataset to find their optimal values. As a result, we will be comparing the accuracy of Wu--Lin--Weng method evaluated with repeated five-fold cross-validation accuracy of a method in family \eqref{eq:family}. We use 20 repetitions.
\item To compare two parameter vectors, we compare their resulting cross-validation accuracy. An alternative approach would be to use  a gradient based optimization for a smooth proxy (e.g. cross-entropy) of empirical accuracy. We opted for the former because the problem is two-dimensional, and we can use grid search instead.
\item Our choice of grid search necessitates a selection of a compact subset of parameter space that will be sampled. We choose to examine points in the simplex defined by  $a_i\geq 0$, for $i=1,2,3$. As Table \ref{tab:step2} shows, this simplex contained in more than half cases a Bayes covariant method matching the performance of Wu--Lin--Weng's method.
\end{itemize}

\begin{figure}[!ht]
\includegraphics{exp2-plot2.pdf}
\caption{Comparison of accuracy for parametric Bayes covariant methods \eqref{eq:family} vs. Wu--Lin--Weng's method}
\label{fig:par-bc}
\end{figure}

The summary statistics are shown in Figure \ref{fig:par-bc}.
We can see that only in minority of cases we found a parametric Bayes covariant coupling method matching the performance of Wu--Lin--Weng's method.

%\subsection{Experiment: }
%
%We determined that among the 1000 three-class problems we have examined, the largest gap between Bayes-covariant parametric methods and Wu--Lin--Weng method occurred for mnist dataset and classes 5,6,9. Figure 
%
%\begin{itemize}
%	\item We searched within a much larger rectangle $D$ containing the simplex $a_i\geq 0$, which in this case is a triangle $T$.
%	\item We used Brier score as a smooth proxy for measuring the quality of multi-class classifier. This score is much smoother than 0-1 score which measures accuracy.
%\end{itemize}
%
%The results are shown in Figure \ref{fig:score}
%
%
%\begin{figure}[!ht]
%	\includegraphics{graphs/exp2-detail.pdf}
%	\caption{Color map indicates the value of Brier score computed for an ensemble of coupling methods on the testing dataset. The vertices triangle indicate the arboreal coupling methods, and the black circle the location of the optimum.}
%	\label{fig:score}
%\end{figure}
%
%We can see that indeed the optimum over the rectangle lies outside of the simplex. Let us compare the resulting accuracy of the optima in $\Delta$ and $D$. The results are shown in Table \ref{tab:step5}.
%
%\input tab-step5.tex

\subsection{Experiment: Effect of change of priors}  \label{sec:exp4}
%\subsection{Wu--Lin--Weng's method}

We have shown that are situations when Wu--Lin--Weng's method outperforms even the best Bayes-covariant parametric methods. In this section we thus ponder the question whether some non-Bayes covariant method would be more resilients to change of priors. We consider two facets to this question.

@@

The results, as expected, detect no change between $\boldsymbol{c}_1, \boldsymbol{c}_2$ for normal coupling method. However, there is no consistent difference in spread between radial and Wu--Lin--Weng's coupling method. Spread for the radial method is lower for 3 datasets (letter, mnist, usps datasets), higher for 3 datasets (dna, satimage, segment) and tied for waveform dataset.

@@

\begin{figure}[!ht]
	\includegraphics{exp4-truth.pdf}
	\caption{Difference in accuracies of $\boldsymbol{c}_2$ and $\boldsymbol{c}_1$ for radial and Wu--Lin--Weng's coupling method.}
	\label{fig:score}
\end{figure}

### 2025-08-14 mdpi.tex
Correct hyphenation of Wu-Lin-Weng

\documentclass[twoside,11pt]{article}

@@

\title{New Bayes covariant coupling methods to cope with change of priors in one-vs-one classification }

@@

\begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
%\blindtext
Label shift is one kind of distribution shift affecting classification models. Generally, one uses transformation inspired by Bayes classifiers to 
adjust predictions of probabilistic classification models. However, there is an ambiguity how to apply this transformation in the case of  probabilistic models built in one-vs-one (also called pairwise) fashion. One-vs-one classification relies on the choice of a coupling method. Some coupling methods -- called Bayes covariant --  are not affected by the ambiguity. Adopting a Bayes covariant method would be the ideal solution to the ambiguity. However, the most commonly used coupling  method of Wu-Lin-Weng is not Bayes covariant. In our study we present elements of theory of coupling methods and exhibit new Bayes covariant coupling methods. One class is formed by arboreal methods. Still more methods can be formed by the way of ensembling. We compare their performance with that of Wu-Lin-Weng's coupling method. In majority of cases, it is possible to find a Bayes covariant method matching performance of Wu-Lin-Weng's method.

\end{abstract}

\begin{keywords}
  distribution shift, label shift, pairwise coupling, one-vs-one classification, Bayes classifier
\end{keywords}

\input intro.tex

@@

\acks{Research on this paper was partially supported by grant VEGA 2/0172/22 Classification using ensembles of neural networks. We thank Muhammad Azeem for helpful discussion during preparation of our paper. }

>>>>

%\documentclass[twoside,11pt]{article}
\documentclass[journal,article,submit,pdftex,moreauthors]{Definitions/mdpi}

@@

\title{On selection of a coupling method to cope with change of priors in one-vs-one classification }

@@

\begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
\blindtext
\end{abstract}

\begin{keywords}
  keyword one, keyword two, keyword three
\end{keywords}

\section{Introduction}

%\blindmathpaper

Distribution shift is an important problem facing machine-learning practitioners  [\cite{zhang2023dive, sugiyama2007covariate}].  One particular way a distribution shift may manifest in classification problems  is when the class priors change between the training set used to train a model, and the data distribution in deployment -- the so called label shift [\cite{vsipka2022hitchhiker,lipton2018detecting}]. To make a concrete example, consider a driver assistance system recognizing traffic signs. Suppose a speed limit sign was maliciously altered to indicate 80mph limit instead of 30mph. To avoid being fooled by the alteration, the system could utilize priors. The distribution of traffic signs will change markedly when the car leaves highway and enters streets in a city. It is very unlikely to encounter 80mph limit in a city setting. By taking into account the city priors, a driver assistance is likely to make the proper classification. In general a good classification system will be able to improve its performance when supplied with a new set of class priors that correspond to the true distribution of classes.

Change of priors can be effectively dealt with in probabilistic classifiers. The basic idea is that Bayes theorem stipulates \emph{exactly} what should happen  to the prediction of the  Bayes classifier when priors change.  In practice we assume that a trained classifier closely approximates the Bayes classifier. Therefore we can make adjustment to our prediction by applying the same transformation of posterior as would happen with the Bayes classifier.

There is a subtle issue with this approach  when handling probabilistic classifiers arising in one-vs-one classification framework.   One-vs-one classification is a two-step framework to approach  multi-class classification. In the first step it reduces the multiclass problem to a series of binary classification problems. In the second step it deduces the multi-class decision by aggregating decisions for the binary problems. Since the procedure is two-step, there is a possibility to apply adjustment of priors either to the probabilistic distributions obtained after the first step, or after the second step. It is not clear what should be the preferred course of action and this problem motivates our paper.

Our approach to the problem is to evaluate practical performance of different methods to aggregate binary predictions. Such methods -- called coupling methods -- first appeared in work of \cite{refregier1991probabilistic}. Many authors followed with alternative proposals [\cite{price1994pairwise, hastie1998classification,  zahorian1999partitioned, wu2004probability, vsuch2015new, vsuch2016bayes}].

We formulate the underlying principles of one-vs-one probabilistic classification in Section 2. In Section 3 we present  different classes of coupling methods. We pay special focus to methods that commute with change of priors -- Bayes covariant methods, since such methods provide an unexpected, and elegant solution to our problem. We introduce new coupling methods by forming weighted ensembles, which we prove are also Bayes covariant when a normalizing condition on weights hold (Proposition \ref{prop:ensemble}).  In Section 4 we summarize our methodology. The results of our experiments are described in Section 5. In Section 6 we summarize our findings and provide practical guidance to solving the problem.

@@

\acks{All acknowledgements go at the end of the paper before appendices and references.
Moreover, you are required to declare funding (financial activities supporting the
submitted work) and competing interests (related financial activities outside the submitted work).
More information about this disclosure can be found on the JMLR website.}
